---
title: "Workshop 4: Research Question #3"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

In this workshop, we first reviewed the code for Research Question #1 and discussed the importance of "commenting" our code (aka leaving notes for ourselves in the future). Although we didn't have time to go through the code for Research Question #2, you can find the code in the workshop2_answers.Rmd on RStudio cloud. 

We spent the majority of our time working through Research Question #3: Is there a relationship between height and weight for basketball players (both male and female)? With this research question, we talked about the concept of a correlation. 

We spent the remaining time working in your groups to discuss ideas for your own research question. 

# Load libraries

As always, the first thing we need to do is load in the libraries (or packages) we want to use

```{r}
library(tidyverse)
```

# Read in data

```{r}
olympics <- read_csv("olympics.csv")
olympics
```


# Research Question 3

Is there a relationship between height and weight for basketball players (both male and female)?

```{r}
olympics %>%
  select(sex, height, weight, sport) %>%
  na.omit() %>%
  filter(sport == "Basketball") %>%
  group_by(sex) %>%
  summarise(correlation = cor(height, weight)) 
```
